### 2014-05-04 chapter2.tex
Updated makefile to give a word count after building PDF. Made some good progress with chapter 2 which is nearly completion. Also broke each chapter out into its own tex file to make it easier to manage. This has given a nice fringe benefit of being able to report word count broken down by chapter.

\chapter{Linked Data}
\label{stateoftheart}

The concept of Linked Data can well be summarised as a representation of
data that follows four principles:\cite{berners2011linked}

\begin{enumerate}
  \item Use URIs to denote things.
  \item Prefer HTTP URIs (i.e. URLs) so that these things can be looked
    up (i.e. \emph{dereferenced}.
  \item When these URIs are dereferenced, provide useful information about
    those things using standards.
  \item Include links to other things (using their URIs) when such
    information is published on the web.
\end{enumerate}

It can be seen as a subset of Semantic Web techniques and it builds upon
the \emph{Resource Description Framework} (RDF)\cite{lassila1999resource},
which is a general-purpose method for describing and modelling information.

In this chapter, I will outline how web authors are encouraged to publish
linked data within their content and how that is already used to enhance
search results in global search engines. I will then introduce how linked
data may be used within an enterprise setting to reduce traditional
integration efforts and then touch on some other aspects of the linked
data domain that could be of interest to the search application.

\section{Publishing Linked Data within Web Pages}
\label{publishing-linked-data}

Whilst there exist multiple formats for publishing RDF purely for machines
(e.g. Turtle\cite{world2014rdf}, N-Triples, JSON-LD\cite{world2014json},
N3, RDF/XML) there are multiple ways to embed semantic or linked data within
an HTML web page:

\begin{itemize}
  \item RDFa\cite{adida2012rdfa} and RDFa Lite\cite{lite2004rdfa}.
  \item DC-HTML defines how to embed the Dublin Core vocabulary within HTML
    meta tags.
  \item Microdata is an HTML standard that allows annotating existing markup
    via additional attributes so that it serves as metadata in addition to
    its purpose as text for humans.
  \item Microformats are another mechanism that embeds semantic metadata
    within existing HTML tags and attributes.
\end{itemize}

Using one or more of these approaches allows a publisher to convey
machine-readable metadata in the same document that is served to humans.
This allows, for example, search engines to index more categorical and
structured information about a page when it is crawled in addition to the
normal processing on the text content. This was the foundation of the
Semantic Web\cite{berners2001semantic} before the term Linked Data
emerged and the aspiration was that widespread adoption would lead to
a comprehensive ``Web of Data'' to rival the rich, so-called
``Web of Documents'' that is the World Wide Web as we know it as a collection
of human-readable pages.

People such as Hepp have argued against adding markup such as RDFa around
the text visible to humans and instead adding hidden, additional markup to
contain semantic information.\cite{hepp2009rdf2rdfa} Hepp argues that
this allows the HTML tree structure to be decoupled from the data structure
of the vocabulary in use, although it does introduce redundancy if all
semantic information is duplicated in both human- and machine-readable forms.

Google have seemingly partially adopted an approach following this principle
-- at least for web authors publishing information about music
events.\cite{googlejsonld} Google are piloting embedding RDF metadata in
\emph{JSON-LD} format within HTML \texttt{<script>} tags to add semantic
information alongside textual data, but without annotating this metadata
around the textual content itself as we would with RDFa, for example.

All of the above approaches can be found ``in the wild'' in pages across
the Web with differing rates of adoption. General-purpose web search
engines can make use of this additional metadata to improve the quality
of the search results in a number of ways. In section~\ref{enhancing-results},
I discuss some of these potential improvements.

\section{Enhancing Search Engine Results}
\label{enhancing-results}

There are at least two key ways in which a search application might
improve the quality of results by making use of semantic metadata as
described in section~\ref{publishing-linked-data}.

One approach is to allow
user queries to be matched against metadata properties in addition
to the text, perhaps with higher precedence. This may lead to a search
engine that seems ``intelligent'' enough not just to retrieve articles
by keywords found therein, but also by keywords that match related
concepts and properties.

For example, a page about a film might be retrieved
by matching names of actors that appeared in the film, even if that actor's
name does not appear in any textual information -- e.g. the synopsis --
that appears on the page. If the author has published semantic metadata
within the the film's page to list all the actors and other contributors,
then our search application can match on this auxiliary information.

Whilst this technique could improve the relevance of the results for a
given query (and consequently improve the precision of the system for that
particular query), a second technique known as \emph{Rich Snippets} improves
results from a usability perspective by aiding users in assessing whether
a result is relevant or not.

Rich Snippets are a distinct improvement 
to the user experience and information architecture
of search applications. A search engine results page (sometimes known as a
\emph{SERP}) may well retrieve almost all relevant results for a query -- and
achieve nearly one hundred percent precision -- but this high performance
is lost on a human user if the results are presented in a form where it is
hard to determine the respective relevant of each result.

\section{Linked Enterprise Data}

In recent years, a growing trend has emerged to use
\emph{Linked Data} in the enterprise to remove the need
for repeated integration projects or data warehousing.
\cite{allemang2010semantic} If we view the indexing processes
of a search application as a variant of data warehousing,
we can start to argue for enterprise-wide publishing of
linked data as a more cost effective means of building search
indexes over bespoke integration efforts.

An important consideration is that BBC Search is largely directing
users to public pages on the BBC website, a large number of which have
already published some of their metadata as linked data. Notably, BBC News,
BBC Sport and BBC Food all use some combination
of microformats, microdata, RDFa, HTML meta tags within their HTML content
pages. Some of this was to improve search results in external search engines
such as Google and some of it ties into Facebook's Open Graph Protocol, which
enables and enhances social media sharing of the pages.

These efforts are not as structured and rich as the more notable uses
of semantic web technologies on the programmes microsites, Nature and BBC
Music \cite{raimond2010use}, but indicate that some level of metadata
can be retrieved
as linked data. In cases where the metadata are rich enough, it might be
possible to avoid direct integration with databases, hidden back end
web services and other structured ``feeds'' typically set up for enterprise
integration.

\section{Reasoning and Inference}