### 2014-05-04 paper.tex

\documentclass[10pt,a4paper]{report}

\usepackage{nag}
\usepackage{mathtools}
\usepackage{natbib}
\usepackage{url}

\title{Proposal to improve BBC Search through the use of Linked Data}
\author{Ross Fenning}

\begin{document}

\maketitle

\include{chapter1}
\include{chapter2}
\include{chapter3}
\include{chapter4}
\include{chapter5}
\include{chapter6}

\bibliographystyle{plain}
\bibliography{bibtex}

\end{document}