### 2014-05-05 chapter2.tex
Drawing a line under chapter 2 for now so I can move on to the important stuff. I've not said anything around inference, but I probably don't need to.

\chapter{Linked Data}
\label{stateoftheart}

The concept of Linked Data can well be summarised as a representation of
data that follows four principles:\cite{berners2011linked}

\begin{enumerate}
  \item Use URIs to denote things.
  \item Prefer HTTP URIs (i.e. URLs) so that these things can be looked
    up (i.e. \emph{dereferenced}.
  \item When these URIs are dereferenced, provide useful information about
    those things using standards.
  \item Include links to other things (using their URIs) when such
    information is published on the web.
\end{enumerate}

It can be seen as a subset of Semantic Web techniques and it builds upon
the \emph{Resource Description Framework} (RDF)\cite{lassila1999resource},
which is a general-purpose method for describing and modelling information.

In this chapter, I will outline how web authors are encouraged to publish
linked data within their content and how that is already used to enhance
search results in global search engines. I will then introduce how linked
data may be used within an enterprise setting to reduce traditional
integration efforts and then touch on some other aspects of the linked
data domain that could be of interest to the search application.

\section{Publishing Linked Data within Web Pages}
\label{publishing-linked-data}

Whilst there exist multiple formats for publishing RDF purely for machines
(e.g. Turtle\cite{world2014rdf}, N-Triples, JSON-LD\cite{world2014json},
N3, RDF/XML) there are multiple ways to embed semantic or linked data within
an HTML web page:

\begin{itemize}
  \item RDFa\cite{adida2012rdfa} and RDFa Lite\cite{lite2004rdfa}.
  \item DC-HTML defines how to embed the Dublin Core vocabulary within HTML
    meta tags.
  \item Microdata is an HTML standard that allows annotating existing markup
    via additional attributes so that it serves as metadata in addition to
    its purpose as text for humans.
  \item Microformats are another mechanism that embeds semantic metadata
    within existing HTML tags and attributes.
\end{itemize}

Using one or more of these approaches allows a publisher to convey
machine-readable metadata in the same document that is served to humans.
This allows, for example, search engines to index more categorical and
structured information about a page when it is crawled in addition to the
normal processing on the text content. This was the foundation of the
Semantic Web\cite{berners2001semantic} before the term Linked Data
emerged and the aspiration was that widespread adoption would lead to
a comprehensive ``Web of Data'' to rival the rich, so-called
``Web of Documents'' that is the World Wide Web as we know it as a collection
of human-readable pages.

People such as Hepp have argued against adding markup such as RDFa around
the text visible to humans and instead adding hidden, additional markup to
contain semantic information.\cite{hepp2009rdf2rdfa} Hepp argues that
this allows the HTML tree structure to be decoupled from the data structure
of the vocabulary in use, although it does introduce redundancy if all
semantic information is duplicated in both human- and machine-readable forms.

Google have seemingly partially adopted an approach following this principle
-- at least for web authors publishing information about music
events.\cite{googlejsonld} Google are piloting embedding RDF metadata in
\emph{JSON-LD} format within HTML \texttt{<script>} tags to add semantic
information alongside textual data, but without annotating this metadata
around the textual content itself as we would with RDFa, for example.

All of the above approaches can be found ``in the wild'' in pages across
the Web with differing rates of adoption. General-purpose web search
engines can make use of this additional metadata to improve the quality
of the search results in a number of ways. In section~\ref{enhancing-results},
I discuss some of these potential improvements.

\section{Enhancing Search Engine Results}
\label{enhancing-results}

There are at least two key ways in which a search application might
improve the quality of results by making use of semantic metadata as
described in section~\ref{publishing-linked-data}.

One approach is to allow
user queries to be matched against metadata properties in addition
to the text, perhaps with higher precedence. This may lead to a search
engine that seems ``intelligent'' enough not just to retrieve articles
by keywords found therein, but also by keywords that match related
concepts and properties.

For example, a page about a film might be retrieved
by matching names of actors that appeared in the film, even if that actor's
name does not appear in any textual information -- e.g. the synopsis --
that appears on the page. If the author has published semantic metadata
within the the film's page to list all the actors and other contributors,
then our search application can match on this auxiliary information.

Whilst this technique could improve the relevance of the results for a
given query (and consequently improve the precision of the system for that
particular query), a second technique known as \emph{Rich Snippets} improves
results from a usability perspective by aiding users in assessing whether
a result is relevant or not.

Rich Snippets are a distinct improvement 
to the user experience and information architecture
of search applications. A search engine results page (sometimes known as a
\emph{SERP}) may well retrieve almost all relevant results for a query -- and
achieve nearly one hundred percent precision -- but this high performance
is lost on a human user if the results are presented in a form where it is
hard to determine the respective relevant of each result.

Evaluation of the effectiveness of search results enhanced with rich snippets
via both explicit and implicit user feedback had been carried out by
Haas et al.\cite{haas2011enhanced} at Microsoft and Yahoo! research. The
results showed a promising preference from 84\% of users for a result
with enhanced information with the most common downside being that the
snippet may sometimes present incorrect information (e.g. showing the wrong
price for a product is less preferable than simply not attempting to present
the price at all). Implicit feedback via click-through rates (CTR) --
a common metric to indicate that users are happy enough with the information
given to click on the link given -- suggested users were more likely to
follow enhanced results.

A higher CTR for semantically-enhanced search results creates some pressure
and incentive on web publishers to include semantic metadata in their web
content if they aspire to attacting visitors via global search engines such
as Google, Bing and Yahoo!.

In section~\ref{linked-enterprise-data} and section~\ref{reasoning}, I will
outline some of the linked data approaches and techniques that relate to
working with data models themselves.

\section{Linked Enterprise Data}
\label{linked-enterprise-data}

\emph{Enterprise integration} is an important technical field within the
discipline of Enterprise Architecture that focuses on interconnecting
systems and allowing data to flow between them. This can be seen as
the practice of breaking down information silos that naturally emerge
in enterprise organisations.\cite{allemang2010semantic}

Architectural patterns have been developed, tested and
implemented in a large number of enterprises to synchronise,
transform and normalise data. Such patterns are notably
summarised by Hohpe and Woolf.\cite{hohpe2004enterprise}

There are difficulties with some of these approaches noted by
Allemang\cite{allemang2010semantic}. Data warehousing,
integration projects and metadata repositories can all
be time-consuming to implement, require repeated updating when
use cases change and at the extreme can be seen as solving the
problem of too many silos by adding another silo.

In recent years, a growing trend has emerged to use
Linked Data in the enterprise to remove the need
for repeated integration projects or data warehousing.
indexes over bespoke integration efforts. This follows from
a major motivation of Linked Data being that it aids in breaking
down information silos. \cite{bizer2009linked} In the
\emph{Linked Enterprise Data} approach, all information
production remains \emph{distributed} but is
\emph{connectable}.\cite{allemang2010semantic}

An organisation that seeks to ensure that all information is
published as Linked Data can learn from examples such as Wikipedia
-- or indeed the web in general -- where individual experts are
encouraged to share knowledge rather than keep it in silos. This
distributed approach recognises that diversity and flux are the
steady state of an agile enterprise. Two sources that happen
to use the same vocabulary are instantly interoperable. Two
sources that use a different vocabulary could well be integrated
simply by creating appropriate mappings between the vocabularies.
It should not be necessary to re-engineer a whole dataset nor
create entirely new systems just for integration.

There are still costs however in additional services, infrastructure
and training in exposing data as linked data throughout the
enterprise.\cite{hyland2010preparing} A practical approach
has been suggested by Davis\cite{davis2011achieving} where enterprises
can expose data via known formats such as Atom\cite{nottingham2005atom}
and RESTful web services. Davis argues that the \texttt{atom:link}
element offers some of the functionality of an RDF triple where
an entity can be linked to other entities via the \texttt{href}
attribute and the nature of that relationship is described by
the \texttt{rel} attribute.

The use of RESTful services and XML formats are very common in
the enterprise -- noting however that many do not follow all tenets
of the REST architectural style. This suggests that businesses
could adopt the pattern suggested by Davis to expose their
data across the organisation using technologies and techniques
with which they are likely familiar. This eschews the full power
of a stack based RDF, OWL and SPARQL, but might be a better
opportunity for businesses that otherwise would not take on a
Linked Data approach at all.  Many proponents of Linked Data would
argue that the important aspect is that the data are exposed and
linked, however that is achieved.

\begin{comment}
If we view the indexing processes
of a search application as a variant of data warehousing,
we can start to argue for enterprise-wide publishing of
linked data as a more cost effective means of building search

An important consideration is that BBC Search is largely directing
users to public pages on the BBC website, a large number of which have
already published some of their metadata as linked data. Notably, BBC News,
BBC Sport and BBC Food all use some combination
of microformats, microdata, RDFa, HTML meta tags within their HTML content
pages. Some of this was to improve search results in external search engines
such as Google and some of it ties into Facebook's Open Graph Protocol, which
enables and enhances social media sharing of the pages.

These efforts are not as structured and rich as the more notable uses
of semantic web technologies on the programmes microsites, Nature and BBC
Music \cite{raimond2010use}, but indicate that some level of metadata
can be retrieved
as linked data. In cases where the metadata are rich enough, it might be
possible to avoid direct integration with databases, hidden back end
web services and other structured ``feeds'' typically set up for enterprise
integration.
\end{comment}