### 2014-05-11 chapter4.1.tex

\section{Linked Data Indexing}

The concept of \emph{Linked Enterprise Data} was introduced and described
in section~\ref{linked-enterprise-data}. In this section, it is proposed
to endeavour to make use of this approach where possible in the
ingest and indexing processes of the BBC Search application.

\begin{figure}
  \begin{center}
\begin{dot2tex}[dot,scale=0.55]
digraph ingest {
  rankdir=TB;

  Programmes [shape=rect,label=iPlayer]
  News [shape=rect]
  Newsround [shape=rect]
  Newsbeat [shape=rect]
  Sport [shape=rect]
  MarketData [shape=rect]
  WorldService [shape=rect]
  BBCFood [shape=rect]
  ProgrammesWebsites [shape=rect,label=Programmes]
  LegacyWebsites [shape=rect,label="Legacy Sites"]

  SearchIndexes [shape=circle]

  Programmes -> URLNotification [label="URLs"]

  News -> CPS
  Newsround -> CPS
  Newsbeat -> CPS
  Sport -> CPS

  CPS -> URLNotification [label="URLs"]

  URLFetch -> "www.bbc.co.uk" [label="HTTP\ GET"]
  "www.bbc.co.uk" -> URLFetch [label="RDFa/Microdata"]

  trans3 [label="Transformation"]
  WorldService -> trans3
  MarketData -> trans3
  trans3 -> IngestAPI [label="Atom"]

  ProgrammesWebsites -> Crawler
  BBCFood -> Crawler
  LegacyWebsites -> Crawler

  Crawler -> URLNotification [label="URLs"]

  URLNotification -> URLFetch [label="RDF"]

  URLFetch -> Enrichment [label="RDF"]
  URLFetch -> Inference [label="RDF"]

  Enrichment -> Inference [label="RDF"]

  Inference -> IngestAPI [label="RDF"]

  IngestAPI -> SearchIndexes

}
\end{dot2tex}
  \end{center}
  \caption{Abstract model of proposed indexing process}
  \label{led-indexing-chain}
\end{figure}

An abstract overview of the flow of content is given in figure~\ref{led-indexing-chain}.
As with the current system depicted in figure~\ref{indexing-chain},
there remains an ``Ingest'' API to which any system may
send content to be indexed if it has already been transformed to
use the semantics required by the search application. This
ensures that integration work done by other teams continues to work
and retains the use case where content producers wish to have full
control of what metadata is sent to the search indexed -- as long as
they are willing to do the integration work.

Bespoke integration in this way can still suffer from maintainability
problems that occur when systems are too tightly coupled. Due to this,
it is architecturally important that the Ingest API continues to follow
the REST architectural style, making appropriate use of Hypermedia to
evolve the API's semantics over time whilst not breaking existing clients.

Where this proposal differs from the existing approach is in the case
of more ``light touch'' integration for teams who are less able or willing
to do bespoke transformation work and increasing coupling to the search
APIs. The goal of this part of the proposal is to reduce the integration
and consequent maintenance work for the search development team.

For these cases, the indexing chain starts from the \emph{URLNotification}
stage to which messages can be sent in
\texttt{text/uri-list}\cite{amundsen2011hypermedia} format. It is expected
that there is some mechanism by which this service can be notified with URLs
of every piece of content that has been created or updated on the BBC website.
This might take the form of the search team's own crawler that is seeking
out pages that have not been found yet or it could be a message-driven
notification process. It is hoped that whilst it might be demanding to
put the onus on a team
to provide transformed content and to integrate against the search system's
APIs, it is a far lighter piece of work to send simple notifications containing
a URL.

The URLNotification service can generate a minimal RDF graph and the
rest of the indexing chain can be seen as a process that incrementally
augments this graph until it is suitable for indexing and has sufficient
information to display search results. It is proposed that these minimal
graphs will take the form of the following example (in JSON-LD serialisation):

\begin{centering}
\begin{lstlisting}[language=json]
[
  {
    "@id": "http://www.bbc.co.uk/news/technology-27346567",
    "http://purl.org/dc/terms/identifier": [
      {
        "@value": "http://www.bbc.co.uk/news/technology-27346567"
      }
    ]
  }
]
\end{lstlisting}
\end{centering}

This single triple, whilst arguably tautological, can serve as a stub
for the series of augmentations further downstream. The next stage,
the \emph{URLFetch} stage, is
the attempt to add semantic metadata found by dereferencing the URL
given. This step should retrieve metadata similar to the graphs depicted
in chapter~\ref{analysis} such as figure~\ref{keepod}. If the identifier
in the stub RDF graph sent to this stage is a URI, but not a URL, then
dereferencing is not possible, but the \emph{URLFetch} stage could
passively send the graph onwards without augmenting to it.

In the URLFetch stage, it might be necessary to apply some web scraping/crawling
techniques to extract the body of text from articles where appropriate.
It was identified in the case of figure~\ref{keepod} that BBC News
was not semantically highlighting the actual text content of the articles, but
a search application is certainly interested in this information.

The URLFetch service can then send the augmented graph onwards to
the \emph{Enrichment} service. The purpose of this stage is to apply
any number of business rules to add further metadata to the graph
that has not been obtainable via the public website or further upstream. It is
in this service that the search team is doing more classical
\emph{enterprise integration} work, which is unavoidable, but confined
to this one service (or collection of microservices).

In the Enrichment service, we might, for example, contact backend
service applications for futher data about the content we are trying to ingest.
This might mean fetching information about a programme from services
behind BBC iPlayer or contacting content management systems for
non-published metadata about articles, videos, etc.

Depending on the nature of the data source, this might allow us to start
making \emph{closed world} assumptions about the properties of content
items. In section~\ref{programmes}, it was noted that the \emph{open world}
assumption of linked data would make it difficult to assert negative
things about the content (e.g. this programme is not available to
watch on iPlayer). If we know that our RDF graph has been enriched
with a totality of information \emph{from source}, then it might be
possible to start making some closed world assumptions with certain predicates.

Following the same pattern as the URLFetch stage, it is valid behaviour
for the Enrichment service to assess that it is unable to provide
any useful, additional metadata and to pass the RDF graph further
downstream untouched. This allows an iterative approach where the first
version of the service can only enrich from one data source and
further data sources are added over time. Content items that cannot be
enriched at the present time can simply pass through unchanged.

The next stage is to run the RDF graph through an \emph{Inference} process.
This is where custom rules are invoked to handle the different
vocabularies that might be in use. For example, the rule shown below
(in N3 format) would allow us to infer that we can
use \texttt{schema:headline} anywhere we have used \texttt{rnews:headline}
or \texttt{og:title}.

\begin{centering}
\begin{lstlisting}[language=ttl]
@prefix schema: <http://schema.org/> .
@prefix rnews: <http://iptc.org/std/rNews/2011-10-07#> .
@prefix og: <http://ogp.me/ns#> .

{ ?x rnews:headline ?y } => { ?x schema:headline ?y } .
{ ?x og:title       ?y } => { ?x schema:headline ?y } .
\end{lstlisting}
\end{centering}

This is useful if the decision is made to use the Schema.org vocabulary as
the standard within the search indexes and application, but upstream
content from BBC News is using \texttt{rnews:headline} instead. It is this
mixed used of vocabularies that was a major driver in historic
enterprise integration efforts. In bespoke integration applications, explicit
transformation would be used to translated  ``news item'' to a
``search item'' including the rule that explicitly extracted the
\texttt{rnews:headline} property and set it on the \texttt{schema:headline}
property on the new ``search item'' being created.

In the proposed approach, the RDF model provides a generic framework where
transformation does not need to occur at any syntactic level and we
can use resuable inference rules to make more blanket statements such as
``Where the title has been set for Facebook's Open Graph, use the same
thing for the title on search results''. This rule can then apply to \emph{all}
content that uses that property without regard to what ``kind'' of content item
it is.

This may sound too open to problems where properties are miused in
unexpected ways and that inferring information in key vocabularies from
ones used by other systems might allow for ``dirty data'' to appear. Such
difficulties with the variable quality of linked data on the wider semantic
web were certainly encountered by Hogan et al.\cite{hogan2011searching} when
building the SWSE: the Semantic Web Search Engine. It is argued however,
that these risks are substantially lower for linked enterprise data as
the domain is far more controlled than the public web. It is also an important
feature of the web (both the document web and the semantic web) that linking and
decentralisation bring overall maintainability and interoperability benefits
that outweigh minor integrity concerns that organisations tradtionally attempt
to prevent with strict, controlled vocabularies, governance and bespoke
software integration.

This proposal argues it is more beneficial to take a more agile approach
of validating some minimum requirements of the information before it
enters the search indexes themselves. Any content that is not thus
being accepted can be flagged so that the rules can be amended. For example,
it might be decided that the \texttt{rnews:headline} property is only
correctly used by BBC News and other content producers use it incorrectly. In
this case, we can update our rules as shown below:

\begin{centering}
\begin{lstlisting}[language=ttl]
@prefix schema: <http://schema.org/> .
@prefix rnews: <http://iptc.org/std/rNews/2011-10-07#> .
@prefix og: <http://ogp.me/ns#> .

{ ?x rnews:headline ?y
. ?x og:site_name "BBC News" } => { ?x schema:headline ?y } .
{ ?x og:title             ?y } => { ?x schema:headline ?y } .
\end{lstlisting}
\end{centering}

Here we add an additional constraint
that we only infer the \texttt{schema:headline} property that we might
ultimately use on the search results page from the \texttt{rnews:headline} if
the content has a value for \texttt{og:site\_name} that indicates it came
from BBC News. This can then be iterated on for future rules that define
different behaviour based on different aspects of the content. These inference
rules can be as simple or as complex as needed.

It is even possible to have
completely different sets of rules invoked based on the types of the content
being indexed. Such an approach should be taken with care so as to prevent
being overly brittle and coupled to the exact structure of the content. More
generic rules allow us to add information based on abstract features of the
metadata that might be present across many ``types'' of content.