### 2014-05-12 chapter5.tex

\chapter{Implementation Plan}

In the previous chapter, some proposals for improvements to
the search application were outlined, but without any order
of importance, business value nor strategy for implementing them.
In this chapter, I will align these proposals within a road map
that describes how and when these proposals could be implemented
over a twelve month period.

With the view to follow an \emph{Agile} methodology, the use
of \emph{Story Mapping} described by Gigante\cite{gigante2013creating}
is likely to be a more adaptable and managable approach than
trying to plan every detail over a year.