### 2014-05-12 chapter4.2.tex

\section{Search Results Page}

The overall information architecture, design and user experience of
the results page are out of scope for this proposal, but it is likely
there could be some improvements in this area, particulary around making
the web page itself responsive\cite{marcotte2010responsive}. A separate
study might consider reviewing the use cases of BBC Search from
the viewpoint of website visitors and consider the relevant
\emph{search modes} and \emph{mode chains}, which could lead to
an information architecture that would drive metadata requirements.

In this proposal, the focus is to propose what \emph{could} be
achieved in an automated way when the principles of linked data
are introduced to the search application. Ultimately, proposed
changes to the search results page should ideally be supported
by valid analysis from a user experience and information
architecture stance.