### 2014-05-13 chapter3.tex

\chapter{Analysis}
\label{analysis}

In this chapter, I will give a deeper analysis of the four
components of the search application identified in
chapter~\ref{intro} and start to highlight areas where linked data
approaches or semantic web technologies could have a r\^ole to play.

\input{chapter3.1}
\input{chapter3.2}
\input{chapter3.3}