### 2014-05-13 chapter4.3.tex

\section{Editor's Choice}

It is proposed there can be
a reduction in some of that manual curation through the use
of linked data and editorial staff can focus on curating
their own linked data. Editor's Choice could evolve into a more
generalised and automated subsystem. A metadata-driven curation system
could use common patterns in the data to promote orders of magnitude more
top links with a fraction of the current effort.

Thus the proposal is to maintain a linked data store in parallel
to the document-oriented search indexes, from which