### 2014-05-15 chapter5.3.tex

\section{Road Map}

The BBC Search application differs somewhat from many other sites in that
the vast majority of the effort, complexity and codebase lives well
below the HTML/CSS/Javascript surface. The web application side follows
a typical stack comprising Javascript/CSS interface, server-side code/templates
to render the pages and a data layer (the search indexes). However, it has been
shown so far that there are orders of magnitude more complexity upstream
of this data layer, but I hope I have shown that linked data technologies
will reduce some of the unnecessary complexity at least.

This has implicates for planning features, stories, tasks, etc. within
an Agile methodology as the first story to display the most basically
functionality on a web page can easily lead to a wealth of engineering
behind the scenes. In this case, there are a lot of systems already in
place, so this road map will attempt to keep features as iterative as possible.

\subsection{Fully-Response Search Results Page}

% Information Architecture
% Link to D. Marland's article

\subsection{Indexing Programmes Microsites}

% PICR

\subsection{Linking to iPlayer}

% Expiry of content - recrawl/refetch

\subsection{Automatic Curation of Programmes Microsites}

% Editor's Choice - Triplestore
% Can start to curate other things ahead of time
% Publishing tool?

\subsection{Improve Results for Crawled Sites}

% CBBC, K&L

\subsection{Showing Next Broadcast for a Programme}

% More rich snippets