### 2014-05-15 chapter5.3.tex
Starting to working through the plan

\section{Road Map}

The BBC Search application differs somewhat from many other sites in that
the vast majority of the effort, complexity and codebase lives well
below the HTML/CSS/Javascript surface. The web application side follows
a typical stack comprising Javascript/CSS interface, server-side code/templates
to render the pages and a data layer (the search indexes). However, it has been
shown so far that there are orders of magnitude more complexity upstream
of this data layer, but I hope I have shown that linked data technologies
will reduce some of the unnecessary complexity at least.

This has implicates for planning features, stories, tasks, etc. within
an Agile methodology as the first story to display the most basically
functionality on a web page can easily lead to a wealth of engineering
behind the scenes. In this case, there are a lot of systems already in
place, so this road map will attempt to keep features as iterative as possible.

\subsection{Fully-Response Search Results Page}

Whilst there are several improvements for end users due to better
curation of suggested links and rich snippets, minor improvements
might best be accompanied with more aesthetic changes to increase
their impact. Also, it is good practice in \emph{Behaviour Driven Development}
(BDD) to frame changes in terms of desired behavioural changes from
the perspective of users of the system. To achieve this, the first
feature to be addressed will be rebuilding a responsive results
page in parallel to the existing one.

When the new, responsive page is viable for public use, a switchover
can happen to move people away from the old page. However, this
switchover does not need to happen until some of the later features
have been developed. Whether the new, responsive page is in
use or not, all features will be defined around improvements to
it.

David Marland stated that the ``core principle in creating a potentially
enormous website that will last forever is to get the information
architecture right in the first place'' in discussion around
building the responsive programmes microsites.\cite{marland2014responsive}
Given that
search is also a potentially enormous website driven by of the
order of millions of data items, it is fitting to map out the
information architecture in this first feature and use that to
drive any domain modelling and vocabularies the search application
needs to use.

Thus this first feature should include some principles of
\emph{Domain-driven Design}\cite{evans2004domain}
in that the search team should identify
a domain model of the search page, the results that appear therein
and a common vocabulary that can describe those items even if
they originated in different systems with differing vocabularies.

A common model across differing BBC content items has been
attempted\cite{fenning2014applicability} at a high level, but
for first iterations of this feature, a simplistic model
such as that shown in figure~\ref{fig:initial-onotology}
would suffice.

\subsection{Indexing Programmes Microsites}

% PICR

\subsection{Linking to iPlayer}

% Expiry of content - recrawl/refetch

\subsection{Automatic Curation of Programmes Microsites}

% Editor's Choice - Triplestore
% Can start to curate other things ahead of time
% Publishing tool?

\subsection{Improve Results for Crawled Sites}

% CBBC, K&L

\subsection{Showing Next Broadcast for a Programme}

% More rich snippets