### 2014-05-16 chapter5.3.tex
More headway with Chapter 5, improved how word counts are done and change RDF graph diagrams to render the DOT via dot2text to make scaling nicer rather than going via PNG.

\section{Road Map}

The BBC Search application differs somewhat from many other sites in that
the vast majority of the effort, complexity and codebase lives well
below the HTML/CSS/Javascript surface. The web application side follows
a typical stack comprising Javascript/CSS interface, server-side code/templates
to render the pages and a data layer (the search indexes). However, it has been
shown so far that there are orders of magnitude more complexity upstream
of this data layer, but I hope I have shown that linked data technologies
will reduce some of the unnecessary complexity at least.

This has implicates for planning features, stories, tasks, etc. within
an Agile methodology as the first story to display the most basically
functionality on a web page can easily lead to a wealth of engineering
behind the scenes. In this case, there are a lot of systems already in
place, so this road map will attempt to keep features as iterative as possible.

\subsection{Fully-Response Search Results Page}

Whilst there are several improvements for end users due to better
curation of suggested links and rich snippets, minor improvements
might best be accompanied with more aesthetic changes to increase
their impact. Also, it is good practice in \emph{Behaviour Driven Development}
(BDD) to frame changes in terms of desired behavioural changes from
the perspective of users of the system. To achieve this, the first
feature to be addressed will be rebuilding a responsive results
page in parallel to the existing one.

When the new, responsive page is viable for public use, a switchover
can happen to move people away from the old page. However, this
switchover does not need to happen until some of the later features
have been developed. Whether the new, responsive page is in
use or not, all features will be defined around improvements to
it.

David Marland stated that the ``core principle in creating a potentially
enormous website that will last forever is to get the information
architecture right in the first place'' in discussion around
building the responsive programmes microsites.\cite{marland2014responsive}
Given that
search is also a potentially enormous website driven by of the
order of millions of data items, it is fitting to map out the
information architecture in this first feature and use that to
drive any domain modelling and vocabularies the search application
needs to use.

Thus this first feature should include some principles of
\emph{Domain-driven Design}\cite{evans2004domain}
in that the search team should identify
a domain model of the search page, the results that appear therein
and a common vocabulary that can describe those items even if
they originated in different systems with differing vocabularies.

\begin{figure}
  \begin{center}
    \begin{dot2tex}[dot,pgf,scale=0.41]
      \input{basic-ontology.dot}
    \end{dot2tex}
  \end{center}
  \caption{Basic ontology modelling the traditional ``title, link, description'' search result}
  \label{fig:initial-ontology}
\end{figure}

A common model across differing BBC content items has been
attempted\cite{fenning2014applicability} at a high level, but
for first iterations of this feature, a simplistic model
such as that shown in figure~\ref{fig:initial-onotology}
would suffice.

The ``Results Page'' subteam can thus focus on initiating
several iterations of user experience design, user testing
and development of the results page and it is expected this
feature will continue to develop in the background as
other features are worked on by other subteams. It is likely
individual stories might add to the domain model even
to make this simple, responsive rebuild of the results page.

The front end application can continue to search the
existing search indexes independant of any work being done
to improve the ingest chain. The importance of developing
this feature first is to ensure the information architecture
and domain models are defined up front. After this point,
the other subteams can proceed with the features below
while the results page developers keep iterating on
a user interaction that evolves through user testing and
feedback.

\subsection{Indexing Programmes Microsites}

The ``Ingest'' subteam needs to ensure the Ingest API can
accept content meeting the basic vocabulary outline in
figure~\ref{fig:initial-onotologu}, validating ones
that do not contain the basic fields and evolve the API. If
the results page developers choose to alter the domain
model based on design iterations and user feedback, then
these changes can propogate back as requirements for the
content ingest.

The team should also evaluate any relevant technologies for the
search indexes and ensure they are searchable by the
developers working on the results page.

Elasticsearch is a standalone search server based on Apache Lucene
that can store, index and retrieve schemaless JSON structures.
The indexing chain is to deal with RDF graphs, for which JSON-LD
is a powerful JSON serialisation that can be sent to Elasticsearch
without any transformation.
A simple prototype of an ingest function written in Python
that receives JSON-LD is shown in listing~\ref{lst:ingest}.

\begin{lstlisting}[language=Python]
from elasticsearch import Elasticsearch
import json

def ingest_jsonld(jsonld_string):
  body = {'jsonld': json.loads(jsonld_string)}
  ElasticSearch().index(index='search', body=body, doc_type='item')
\label{lst:ingest}
\end{lstlisting}

The focus for the ingest work should then be around setting
up mappings in Elasticsearch (if the prototype proves promising)
to arrange indexes to enable sensible retrieval behaviour.

% PICR

\subsection{Linking to iPlayer}

% Expiry of content - recrawl/refetch

\subsection{Automatic Curation of Programmes Microsites}

% Editor's Choice - Triplestore
% Can start to curate other things ahead of time
% Publishing tool?

\subsection{Improve Results for Crawled Sites}

% CBBC, K&L

\subsection{Showing Next Broadcast for a Programme}

% More rich snippets