### 2014-05-17 chapter5.4.tex
Added a wrap-up to chapter 5.

\section{Future Iterations}

The continuing strategy is to move all existing indexing systems into
this RDF-based indexing chain, but also to drive the changes with
tangible improvements visible to users on the results page itself.

In following agile principles, it is remaining out of scope for this
proposal to suggest features beyond those already presented. At a time
where most of these features are implemented, the problem space and
landscape may have changed enough that the road map presented would
have to react to these changes. It is possible that this reaction to
change may have to occur before the described features are fully implemented.